---
title: "EZlab_Vignette"
output: rmarkdown::html_vignette
vignette: >
  %\VignetteIndexEntry{EZlab_Vignette}
  %\VignetteEngine{knitr::rmarkdown}
  %\VignetteEncoding{UTF-8}
---

```{r, include = FALSE}
knitr::opts_chunk$set(
  collapse = TRUE,
  comment = "#>"
)
```
The EZlab functions are created to simplify data frame manipulation for large data sets such as proteomics. 

Install required packages.
```{r}
library(tidyverse)
library(matrixStats)
library(reshape2)
```
Call EZlab package
```{r setup}
library(EZlab)
```
Rename raw_data as df
```{r}
df = raw_data
```
Compute mean and std for each sample. 
```{r}
df <- EZmeanAndstd(df, c('SAMPLE.1', 'SAMPLE.1.1', 'SAMPLE.1.2'), "SAMPLE1")
df <- EZmeanAndstd(df, c('SAMPLE.2', 'SAMPLE.2.1', 'SAMPLE.2.2'), "SAMPLE2")
df <- EZmeanAndstd(df, c('SAMPLE.3', 'SAMPLE.3.1', 'SAMPLE.3.2'), "SAMPLE3")
df <- EZmeanAndstd(df, c('SAMPLE.4', 'SAMPLE.4.1', 'SAMPLE.4.2'), "SAMPLE4")
df <- EZmeanAndstd(df, c('SAMPLE.5', 'SAMPLE.5.1', 'SAMPLE.5.2'), "SAMPLE5")
```
Compute fold change for each sample.
```{r}
df <- EZfoldChange(df, "SAMPLE1")
df <- EZfoldChange(df, "SAMPLE2")
df <- EZfoldChange(df, "SAMPLE3")
df <- EZfoldChange(df, "SAMPLE4")
df <- EZfoldChange(df, "SAMPLE5")
```
Run filter on all samples. 
```{r}
df <- EZfilter(df, "SAMPLE1")
df <- EZfilter(df, "SAMPLE2")
df <- EZfilter(df, "SAMPLE3")
df <- EZfilter(df, "SAMPLE4")
df <- EZfilter(df, "SAMPLE5")
```
Delete genes that do not meet the filter criteria. Here, we are deleting genes in which the "COVAR" value is greater than 3 and only keeping values with "COVAR" less than and equal to 3. 
```{r}
df <- subset(df, COVAR <= 3)
```
Create new data frame with log2 values only.
```{r}
df1 <- select(df,ID,LOG2_SAMPLE1,LOG2_SAMPLE2,LOG2_SAMPLE3,LOG2_SAMPLE4,LOG2_SAMPLE5)
```
Use melt function from reshape2 package to create a column for gene ID matched with its log2 value. This data frame is used to generate a heat map. Ignore errors at this step. 
```{r}
melt_df <- melt(df1)
```
Generate heat map. 
```{r}
EZPlot <- function(data, x, y) {
  x <- enquo(x)
  y <- enquo(y)
  ggplot(melt_df, aes(!!x, !!y)) +
    geom_tile(aes(fill = value), colour = "white") +
    scale_fill_gradient(low = "white", high = "blue")
}
EZPlot(melt_df, variable, ID)
```